Rohan Fray IS608 Final
========================================================


```{r}
library("ggplot2")
library("reshape2")
library("stringr")
```

```{r}
Degrees.earned.by.Level.and.Sex <- read.csv("data/Degrees earned by Level and Sex.csv")
```

We melt Degrees.earned.by.Level.and.Sex by Year to get
```{r}
DegNorm = melt(Degrees.earned.by.Level.and.Sex, id.vars = "YearEnding")
DegNorm$value = as.numeric(gsub(",","",DegNorm$value))
head(DegNorm)
```

We can then add another variable to the dataframe to tell us the difference between male and female
```{r}
DegNorm$MorF <- ifelse(grepl("Male",DegNorm$variable),"Male",(ifelse(grepl("Female",DegNorm$variable),"Female","NA")))
tail(DegNorm)
```

We then categorize the Degree Types
```{r}
DegNorm$variable = as.character(DegNorm$variable)
DegNorm$DegType=str_split_fixed(DegNorm$variable,"[.]",2)[,1]
DegNorm[which(DegNorm$DegType == "First"),]$DegType = "First \nProfessional"
tail(DegNorm)
```

And then subset the data for only the five degree types and to take out the standalone years of 1950 and 1960
```{r}
DegNormSub = DegNorm[which(DegNorm$YearEnding>1960 & DegNorm$DegType != "All"& DegNorm$DegType != "Alldegrees"),]
DegNormSub$DegType <-factor(DegNormSub$DegType, levels = c("Associates","Bachelor","Masters","First \nProfessional","Doctoral"))
```

We are then ready for the first graph, detailing the breakdown of degrees by gender and degree type:
NB: that the scales are not the same for all the degree types
```{r fig.width = 25, fig.height=20}
a <- ggplot(DegNormSub,aes(x=YearEnding, y=value, group = MorF, fill = MorF)) + geom_bar(position="dodge",stat="identity") + labs(x="Year", y = "Number of Degrees in thousands", title = "Breakdown of Degrees by Gender and Degree Type") + theme(plot.title = element_text(size = rel(4)), legend.position = "right", legend.background = element_rect(color = "black"), strip.text.y = element_text(color = "grey33", angle = 45, size = rel(3),face = "italic"), strip.background = element_rect(color = "white", fill = "white"), legend.text = element_text(size = rel(2)), legend.title = element_text(size = rel(2)),axis.text.x = element_text(size = rel(2), color = "grey33",vjust = 1.5), axis.text.y = element_text(size=rel(2), color = "grey33"), axis.title.x = element_text(size=rel(2.5),face = "italic",vjust=1), axis.title.y = element_text(size=rel(2.5),face="italic",vjust =0.3)) 

b <- a + scale_fill_brewer(palette= "Set1",name = "Gender")

c <- b + facet_grid(DegType~., scales = "free_y") 

theme_update(axis.text.x = element_text(angle = 45, hjust = 1), panel.grid.major = element_line(colour = "grey90"), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.ticks = element_blank(), legend.position = "none")

c
```
We see that as Female Degree holders are on a steady increase throughout the years for all degree types, for Males some degree types have either remained the same or have not grown as quickly as Females.


Now we take a look at some finances of higher education

```{r}
Federal.Student.Aid...Funds.Utilized.in.millions <- read.csv("data/Federal Student Aid - Funds Utilized in millions.csv",na.strings="",strip.white=TRUE)
Private.Tuition <-read.csv("data/Private-Tuition.csv",strip.white=TRUE)
Private.FinancialSupport <- read.csv("data/Private-FinancialSupport.csv",strip.white=TRUE)
Private.Salaries <- read.csv("data/Private-Salaries.csv",strip.white=TRUE)

#for simplicity, using private institution data

#first melt where needed and clean data as well as change column names in order to rbind later
#we also change all 4 dataframes to have three columns (Year, attribute, and value)
FedStudentAid = melt(Federal.Student.Aid...Funds.Utilized.in.millions, id.vars = "Type.of.assistance")
FedStudentAid$value = as.numeric(gsub(",","",FedStudentAid$value))
FedStudentAid$variable = gsub("X","",FedStudentAid$variable)
names(FedStudentAid)[names(FedStudentAid)=="variable"]<-"Year"
names(FedStudentAid)[names(FedStudentAid)=="Type.of.assistance"]<-"attribute"

PrivTuition = Private.Tuition
PrivTuition$Private.Tuition = as.numeric(gsub(",","",PrivTuition$Private.Tuition))
PrivTuition$attribute = "Private Tuition"
names(PrivTuition)[names(PrivTuition)=="Private.Tuition"]<-"value"

PrivFinSupp = melt(Private.FinancialSupport, id.vars = "Item")
PrivFinSupp$value = as.numeric(gsub(",","",PrivFinSupp$value))
PrivFinSupp$variable = gsub("X","",PrivFinSupp$variable)
PrivFinSupp$Item = gsub("[..]","",PrivFinSupp$Item)
names(PrivFinSupp)[names(PrivFinSupp)=="variable"]<-"Year"
names(PrivFinSupp)[names(PrivFinSupp)=="Item"]<-"attribute"

PrivSalaries = melt(Private.Salaries, id.vars = "TYPE.OF.CONTROL.AND.ACADEMIC.RANK")
PrivSalaries$value = as.numeric(gsub(",","",PrivSalaries$value))
PrivSalaries$value = PrivSalaries$value * 1000 #as the salary is in thousands
PrivSalaries$variable = gsub("X","",PrivSalaries$variable)
names(PrivSalaries)[names(PrivSalaries)=="variable"]<-"Year"
names(PrivSalaries)[names(PrivSalaries)=="TYPE.OF.CONTROL.AND.ACADEMIC.RANK"]<-"attribute"

#then we combine into one dataframe by year
data<-rbind(FedStudentAid,PrivTuition)
data<-rbind(data,PrivFinSupp)
data<-rbind(data,PrivSalaries)
```

```{r}
#We then take the subset of the data that we are interested in for the years 1980 onwards
#Federal Direct Student Loans (in millions of dollars)
#Private Tuition
#Salaries for Private School Professors
#Financial Support (in millions of dollars)
dsub <- data[which(data$attribute == "Federal Direct Student Loan (FDSL)" | data$attribute == "Private Tuition" | data$attribute == "Private 4-year institutions" | data$attribute == "Professor"),]
dsub <- dsub[which(dsub$Year >= 1980),]
dsub$Year <- as.integer(dsub$Year)
dsub$attribute <- as.character(dsub$attribute)
#we will rename our attributes and then factor and re-level to make plotting easier
dsub[which(dsub$attribute == "Private Tuition"), ]$attribute = "Private Tuition \nCost"
dsub[which(dsub$attribute == "Federal Direct Student Loan (FDSL)"), ]$attribute = "Federal Direct \nStudent Loan \n(in millions)"
dsub[which(dsub$attribute == "Private 4-year institutions"), ]$attribute = "Private Financial Support \n(in millions)"
dsub[which(dsub$attribute == "Professor"), ]$attribute = "Professor Salaries \n(Private Institutions)"
dsub$attribute <- factor(dsub$attribute, levels = c("Federal Direct \nStudent Loan \n(in millions)","Private Financial Support \n(in millions)","Professor Salaries \n(Private Institutions)","Private Tuition \nCost"))
```

If we now wanted to take a look at 10 random samples of our dataset
```{r}
dsub[sample(nrow(dsub),10),]
```

We are now ready for the graph that shows how salaries and tuition have been increasing, and the massive spike in Federal direct student loans.  Note that the scales are not the same.
```{r fig.width = 25, fig.height=20}
ggplot(dsub,aes(x=Year,y=value,fill=attribute)) + geom_bar(position = "dodge",stat="identity") + facet_grid(attribute~., scales = "free_y") +labs(x="Year",y="Dollar amount",title="Financial Educational data over the Years")+theme(plot.title = element_text(size = rel(4)), strip.text.y = element_text(color = "grey33", angle = 45, size = rel(2.5),face = "italic"), strip.background = element_rect(color = "white", fill = "white"), axis.text.x = element_text(size = rel(2), color = "grey33",vjust = 1.5), axis.text.y = element_text(size=rel(2), color = "grey33"), axis.title.x = element_text(size=rel(2.5),face = "italic",vjust=1), axis.title.y = element_text(size=rel(2.5),face="italic",vjust =0.3)) +scale_fill_brewer(palette= "Set2")
```